---
title: "Tracking data 2019 and 2020"
output: html_notebook
---

Must run visualizingtrackingdata.Rmd first!
Merge tracking, plays, games, and PFF data for 2019.

Merge tracking data for 2019 with plays and games data. 

```{r}
merged.games.plays.tracking2019 <- tracking2019 %>% inner_join(games) %>% inner_join(plays) %>% inner_join(PFFScoutingData)

```
Filter data to include only touchback and return results.

```{r}
merged.games.plays.tracking2019 <- filter(merged.games.plays.tracking2019, specialTeamsResult == "Return" | specialTeamsResult == "Touchback")
```

Transform data so that all plays are going in the same direction.

```{r}
merged.games.plays.tracking2019 <- merged.games.plays.tracking2019 %>%
                mutate(x = ifelse(playDirection == "left", 120-x, x),
                       y = ifelse(playDirection == "left", 160/3 - y, y))
```


Filter tracking data to include only the frames in which a catch is made.

```{r}
merged.games.plays.tracking2019 <- filter(merged.games.plays.tracking2019, event == "kick_received")
```


Merge tracking data for 2020 with plays and games data. 

```{r}
merged.games.plays.tracking2020 <- tracking2020 %>% inner_join(games) %>% inner_join(plays) %>% inner_join(PFFScoutingData)

```
Filter data to include only touchback and return results.

```{r}
merged.games.plays.tracking2020 <- filter(merged.games.plays.tracking2020, specialTeamsResult == "Return" | specialTeamsResult == "Touchback")
```

Transform data so that all plays are going in the same direction.

```{r}
merged.games.plays.tracking2020 <- merged.games.plays.tracking2020 %>%
                mutate(x = ifelse(playDirection == "left", 120-x, x),
                       y = ifelse(playDirection == "left", 160/3 - y, y))
```


Filter tracking data to include only the frames in which a catch is made.

```{r}
merged.games.plays.tracking2020 <- filter(merged.games.plays.tracking2020, event == "kick_received")
```


Merge all three data sets together and export to CSV. 

```{r}
merged.games.plays.tracking.all <- rbind(merged.games.plays.tracking2018, merged.games.plays.tracking2019, merged.games.plays.tracking2020)
write.csv(merged.games.plays.tracking.all,"~/Documents/2022bdb/mergedgamesplaystrackingall.csv", row.names = FALSE)
```

Determine what columns can be dropped.

```{r}
unique(merged.games.plays.tracking.all$kickBlockerId)
unique(merged.games.plays.tracking.all$kickContactType)
unique(merged.games.plays.tracking.all$passResult)
unique(merged.games.plays.tracking.all$snapDetail)
unique(merged.games.plays.tracking.all$snapTime)
unique(merged.games.plays.tracking.all$operationTime)
unique(merged.games.plays.tracking.all$gunners)
unique(merged.games.plays.tracking.all$puntRushers)
unique(merged.games.plays.tracking.all$vises)
```
Remove unneeded columns.

```{r}
merged.games.plays.tracking.all = subset(merged.games.plays.tracking.all, select = -c(kickBlockerId,passResult) )
```

```{r}
merged.games.plays.tracking.all = subset(merged.games.plays.tracking.all, select = -snapDetail )
```