---
title: "Negative Binomial Regression"
output: html_notebook
---

Since the target variable is overdispersed, and can be interpreted as a count variable (since yardage is only provided in whole numbers), we use negative binomial regression. We will use repeated k-folds cross validation to fit the model.

Load required libraries.

```{r}
library(caret)
library(MASS)
library(Hmisc)
```

Impute hangTime for missing values.

```{r}
final.data.returns$imputed_hangTime <- with(final.data.returns, impute(hangTime, mean))
final.data.returns.imputed <- subset(final.data.returns, select = -hangTime)
```

Drop receivingTeamWithin5.

```{r}
final.data.returns.imputed <- subset(final.data.returns.imputed, select = -receivingTeamWithin5)
```

Drop netCoverage.

```{r}
final.data.returns.imputed2 <- subset(final.data.returns.imputed, select = -netCoverage)
```

Determine whether any values are missing.

```{r}
final.data.returns.imputed %>%
  summarise_all(funs(sum(is.na(.))))
```


Define training control.
```{r}
train_control <- trainControl(method="cv", number=10)
```

Train the model.

```{r}
# train the model
NBmodel <- train(yardsTillTouchdown ~ kickLength + speedOfClosestCoverage + 
    distanceToClosestCoverage + kickTeamWithin5 + 
    imputed_hangTime, data=final.data.returns.imputed2, trControl=train_control, method="glm.nb")
# summarize results
print(NBmodel)
```

```{r}
NBModel1 <- glm.nb(yardsTillTouchdown ~ ., data = final.data.returns.imputed)
summary(NBModel1)
```

```{r}
NBModel2 <- update(NBModel1, .~. - receivingTeamWithin5)
summary(NBModel2)
```
```{r}
NBModel3 <- update(NBModel2, .~. + receivingTeamWithin5:netCoverage)
summary(NBModel3)
```

```{r}
NBModel4 <- update(NBModel2, .~. - netCoverage)
summary(NBModel4)
```

```{r}
# train the model
NBmodelCV2 <- train(yardsTillTouchdown ~ kickLength + speedOfClosestCoverage + 
    distanceToClosestCoverage + kickTeamWithin5 + imputed_hangTime, data=final.data.returns.imputed, trControl=train_control, method="glm.nb", start =  c(4, 0.1, 0.1, 0.1, 0.1, 0.1))
# summarize results
print(NBmodelCV2)
```

```{r}
NBmodelCV2
```
```{r}
NBmodelCV2$finalModel
```

```{r}
NBmodelCV2$results
```
Impute missing hangTime values in touchback data.

```{r}
final.data.touchbacks$imputed_hangTime <- with(final.data.touchbacks, impute(hangTime, mean))
final.data.touchbacks.imputed <- subset(final.data.touchbacks, select = -c(hangTime, receivingTeamWithin5, netCoverage))
```

Use k-fold validated model to calculate 

```{r}
final.data.returns.imputed2$predictedPosition <- predict(NBmodelCV2, final.data.returns.imputed2)
```


```{r}
plot(final.data.returns.imputed2$yardsTillTouchdown, final.data.returns.imputed2$predictedPosition, xlab = "Yards till Touchdown", ylab = "Model Prediction", main = "Comparing Return Data with Regression Prediction")
abline(a=0, b=1, col="blue")

```
Find residuals.

```{r}
standard_res <- rstandard(NBmodelCV2$finalModel)
```

```{r}
summary(standard_res)
```

Merge standardized residuals with rest of data set.

```{r}
final.data.returns.imputed3 <- cbind(final.data.returns.imputed2, standard_res)
```


```{r}
plot(final.data.returns.imputed3$yardsTillTouchdown, standard_res, ylab='Standardized Residuals', xlab='Yards till Touchdown', main="Kickoff Return Data")
abline(0,0, col="blue")
```


```{r}
final.data.touchbacks.imputed$predictedPosition <- predict(NBmodelCV2, final.data.touchbacks.imputed)
```

Bind prediction data frames together and merge with original dataframe.

```{r}
prediction.data <- rbind(final.data.returns.imputed2, final.data.touchbacks.imputed)
final.data.with.predictions <- left_join(merged.all.distance.final.closestcoverage, prediction.data, by = c("gameId", "playId", "kickLength", "speedOfClosestCoverage", "distanceToClosestCoverage", "kickTeamWithin5", "yardsTillTouchdown"), copy = FALSE,
  suffix = c(".1", ".2"),
  keep = FALSE)
```

Export dataframe as CSV.

```{r}
write.csv(final.data.with.predictions,"~/Documents/2022bdb/final.data.with.predictions.csv", row.names = FALSE)
```

