---
title: "Model fitting"
output: html_notebook
---

Load packages.

```{r}
library(olsrr)
```


Candidate 1: Linear regression

Explanatory variables:
week
quarter (ideally will eliminate this and roll it into gameTimeRemaining, still troubleshooting code)
kickLength
hangTime
kickDirectionActual
kickoffReturnFormation
speedOfClosestCoverage
distanceToClosestCoverage
netCoverage
scoreDifferentialReceivingTeam
gameClockDecimal

Dependent variable: 
yardsTillTouchdown

```{r}
linearModel1 <- lm(yardsTillTouchdown ~ ., data = final.data.returns)
k <- ols_step_backward_p(linearModel1, details = TRUE)
plot(k)
linearModel1
```

```{r}
ols_step_backward_p(linearModel1, details = TRUE)
```






Update the second model. Subtract the variable with the highest p-value.

```{r}
LinearModel3 = update(LinearModel2, . ~ . - kickoffReturnFormation)
summary(LinearModel3)
anova(LinearModel2, LinearModel3)
```

We have a problem! Several of our records are missing hangTime and/or kickoffReturnFormation. We must impute these in order to have all records included, OR just delete all records with missing info. How many are there?

```{r}
sum(is.na(final.data.returns$kickoffReturnFormation))
sum(is.na(final.data.returns$hangTime))

```
12 records are missing kickoffReturnFormation. 274 are missing hangTime. 

Impute missing values using mi package.

```{r}
library(Hmisc)
```
```{r}
final.data.returns$imputed_hangTime <- with(final.data.returns, impute(hangTime, mean))

```

```{r}
library(epiDisplay)
```


```{r}
tab1(final.data.returns$kickoffReturnFormation, sort.group = "decreasing", cum.percent = TRUE)
```
Replace NA in kickoffReturnFormation with "unknown".

```{r}
final.data.returns$kickoffReturnFormation <- final.data.returns$kickoffReturnFormation %>% replace_na("unknown")
```

Replace NA in kickDirectionActual with "unknown".

```{r}
final.data.returns$kickDirectionActual <- final.data.returns$kickDirectionActual %>% replace_na("unknown")
```


Create new data set and delete hangTime column.

```{r}
final.data.returns.imputed <- subset(final.data.returns, select = -hangTime)
```


Linear model 2:

```{r}
LinearModel2 <- lm(yardsTillTouchdown ~ ., data = final.data.returns.imputed)
m <- ols_step_backward_aic(LinearModel2)
m
plot(m)
summary(LinearModel2)
```

Update by subtracting variables.

```{r}
LinearModel3 = update(LinearModel2, . ~ . - kickoffReturnFormation)
summary(LinearModel3)
anova(LinearModel2, LinearModel3)
```

Update by subtracting variables.

```{r}
LinearModel4 = update(LinearModel3, . ~ . - week)
summary(LinearModel4)
anova(LinearModel3, LinearModel4)
```

Update by subtracting variables.

```{r}
LinearModel5 = update(LinearModel4, . ~ . - gameClockDecimal)
summary(LinearModel5)
anova(LinearModel4, LinearModel5)
```

Update by subtracting variables.

```{r}
LinearModel6 = update(LinearModel5, . ~ . - scoreDifferentialReceivingTeam)
summary(LinearModel6)
anova(LinearModel5, LinearModel6)
```
Update by subtracting variables.

```{r}
LinearModel7 = update(LinearModel6, . ~ . - quarter)
summary(LinearModel7)
anova(LinearModel6, LinearModel7)
```

Update by subtracting variables.

```{r}
LinearModel8 = update(LinearModel7, . ~ . - kickDirectionActual)
summary(LinearModel8)
anova(LinearModel7, LinearModel8)
```

