---
title: "Summarizing play data"
output: html_notebook
---

Count number of plays of each type and each result.

```{r}
plays %>%
  group_by(specialTeamsResult, specialTeamsPlayType) %>%
  summarize(n())
```

Count number of kickoffs and punts.

```{r}
plays %>%
  group_by(specialTeamsPlayType) %>%
  summarize(n())
```


Filter plays to include only plays with results of returns or touchbacks.


```{r}
plays.filtered <- filter(plays, specialTeamsResult == "Return" | specialTeamsResult == "Touchback")
```

Filter plays to include only kickoff plays.

```{r}
plays.filtered.kickoffs <- filter(plays.filtered, specialTeamsPlayType == "Kickoff")
```


How many plays had penalties?

```{r}
plays.filtered.kickoffs %>%
  group_by(penaltyCodes) %>%
  summarize(n())
```
Summarize absoluteYardlineNumber for kick received.

```{r}
plays.filtered.kickoffs %>%
  group_by(specialTeamsResult) %>%
  summarize(absoluteYardlineNumber)

summary(plays.filtered.kickoffs$absoluteYardlineNumber)
absYardlineNumHist <- hist(plays.filtered.kickoffs$absoluteYardlineNumber)
absYardlineNumHist
```
Filter plays data to find touchbacks with a returner ID.

```{r}
plays.filtered.touchbacks <- filter(plays.filtered, specialTeamsResult == "Touchback")
```

No touchbacks have a returner ID. 

Summarize absoluteYardLineResult by play direction.

```{r}
merged.games.plays.tracking.all2 %>% group_by(playDirection, absoluteYardlineNumber) %>% summarize(n())

```

Filter plays of interest based on play direction and create histograms.

```{r}
merged.all.distance.leftplays <- filter(merged.all.distance.final.closestcoverage, playDirection == "left")
merged.all.distance.rightplays <- filter(merged.all.distance.final.closestcoverage, playDirection == "right")
LeftAbsYardlineNumHist <- hist(merged.all.distance.leftplays$absoluteYardlineNumber)
RightAbsYardlineNumHist <- hist(merged.all.distance.rightplays$absoluteYardlineNumber)
LeftAbsYardlineNumHist
RightAbsYardlineNumHist
```
Create histogram of net coverage.

```{r}
NetCoverageHist <- hist(merged.all.distance.final.closestcoverage$netCoverage, breaks=seq(-4.5,4.5,by=1))
NetCoverageHist
```
Filter play data by kickoffs and returns vs touchbacks.

```{r}
plays.kickoffs.touchbacks <- filter(plays.filtered.kickoffs, specialTeamsResult == "Touchback")
plays.kickoffs.returns <- filter(plays.filtered.kickoffs, specialTeamsResult == "Return")
```

Find kickoff return plays that had returns for touchdowns.

```{r}
plays.filtered.kickoffs.touchdowns <- filter(plays.filtered.kickoffs, kickReturnYardage > 90)
```

```{r}
unique(final.data.returns.imputed$kickDirectionActual)
```

