---
title: "Visualizing Player Tracking Data"
output: html_notebook
---

Load necessary packages.
```{r}
library(tidyverse) 
library(lubridate)
library(beeswarm)  
library(gganimate) 
library(ggridges)  
library(tidyr)  
library(caret)
library(MASS)
library(Hmisc)
```

Import data.

```{r}
tracking2020 <- read.csv("~/Documents/2022bdb/nfl-big-data-bowl-2022/tracking2020.csv")
tracking2019 <- read.csv("~/Documents/2022bdb/nfl-big-data-bowl-2022/tracking2019.csv")
tracking2018 <- read.csv("~/Documents/2022bdb/nfl-big-data-bowl-2022/tracking2018.csv")
plays <- read.csv("~/Documents/2022bdb/nfl-big-data-bowl-2022/plays.csv")
players <- read.csv("~/Documents/2022bdb/nfl-big-data-bowl-2022/players.csv")
games <- read.csv("~/Documents/2022bdb/nfl-big-data-bowl-2022/games.csv")
```

```{r}
PFFScoutingData <- read.csv("~/Documents/2022bdb/nfl-big-data-bowl-2022/PFFScoutingData.csv")
```


Merge tracking data for 2018 with plays and games data. 

```{r}
merged.games.plays.tracking2018 <- tracking2018 %>% inner_join(games) %>% inner_join(plays) %>% inner_join(PFFScoutingData)

```
Filter data to include only touchback and return results.

```{r}
merged.games.plays.tracking2018 <- filter(merged.games.plays.tracking2018, specialTeamsResult == "Return" | specialTeamsResult == "Touchback")
```

Transform data so that all plays are going in the same direction.

```{r}
merged.games.plays.tracking2018 <- merged.games.plays.tracking2018 %>%
                mutate(x = ifelse(playDirection == "left", 120-x, x),
                       y = ifelse(playDirection == "left", 160/3 - y, y))
```
(Note: In this step, we should have also transformed the variable absoluteYardlineNumber in the same manner that x was transformed. This was an oversight and was corrected later in the notebook distancebetweenplayers.Rmd.)

Filter tracking data to include only the frames in which a catch is made.

```{r}
merged.games.plays.tracking2018 <- filter(merged.games.plays.tracking2018, event == "kick_received")
```

Choose a single play. 

```{r}
merged.games.plays.tracking2018play36 <- filter(merged.games.plays.tracking2018, playId == 36 & gameId == 2018123000)
```

Visualize catch. 


```{r}
## General field boundaries
xmin <- 0
xmax <- 160/3
hash.right <- 38.35
hash.left <- 12
hash.width <- 3.3


## Specific boundaries for a given play
ymin <- max(round(min(merged.games.plays.tracking2018play36$x, na.rm = TRUE) - 10, -1), 0)
ymax <- min(round(max(merged.games.plays.tracking2018play36$x, na.rm = TRUE) + 10, -1), 120)
df.hash <- expand.grid(x = c(0, 23.36667, 29.96667, xmax), y = (10:110))
df.hash <- df.hash %>% filter(!(floor(y %% 5) == 0))
df.hash <- df.hash %>% filter(y < ymax, y > ymin)

animate.play <- ggplot() +
  scale_size_manual(values = c(6, 4, 6), guide = FALSE) + 
  scale_shape_manual(values = c(21, 16, 21), guide = FALSE) +
  scale_fill_manual(values = c("#e31837", "#654321", "#002244"), guide = FALSE) + 
  scale_colour_manual(values = c("black", "#654321", "#c60c30"), guide = FALSE) + 
  annotate("text", x = df.hash$x[df.hash$x < 55/2], 
           y = df.hash$y[df.hash$x < 55/2], label = "_", hjust = 0, vjust = -0.2) + 
  annotate("text", x = df.hash$x[df.hash$x > 55/2], 
           y = df.hash$y[df.hash$x > 55/2], label = "_", hjust = 1, vjust = -0.2) + 
  annotate("segment", x = xmin, 
           y = seq(max(10, ymin), min(ymax, 110), by = 5), 
           xend =  xmax, 
           yend = seq(max(10, ymin), min(ymax, 110), by = 5)) + 
  annotate("text", x = rep(hash.left, 11), y = seq(10, 110, by = 10), 
                    label = c("G   ", seq(10, 50, by = 10), rev(seq(10, 40, by = 10)), "   G"), 
                    angle = 270, size = 4) + 
  annotate("text", x = rep((xmax - hash.left), 11), y = seq(10, 110, by = 10), 
           label = c("   G", seq(10, 50, by = 10), rev(seq(10, 40, by = 10)), "G   "), 
           angle = 90, size = 4) + 
  annotate("segment", x = c(xmin, xmin, xmax, xmax), 
           y = c(ymin, ymax, ymax, ymin), 
           xend = c(xmin, xmax, xmax, xmin), 
           yend = c(ymax, ymax, ymin, ymin), colour = "black") + 
  geom_point(data = merged.games.plays.tracking2018play36, aes(x = (xmax-y), y = x, shape = team,
                                 fill = team, group = nflId, size = team, colour = team), alpha = 0.7) + 
  geom_text(data = merged.games.plays.tracking2018play36, aes(x = (xmax-y), y = x, label = jerseyNumber), colour = "white", 
            vjust = 0.36, size = 3.5) + 
  ylim(ymin, ymax) + 
  coord_fixed()
```

```{r}
animate.play
```

Choose a second play (this is a touchback).

```{r}
merged.games.plays.tracking2018play392 <- filter(merged.games.plays.tracking2018, playId == 392
 & gameId == 2018123000)
```

Visualize the play.


```{r}
## General field boundaries
xmin <- 0
xmax <- 160/3
hash.right <- 38.35
hash.left <- 12
hash.width <- 3.3


## Specific boundaries for a given play
ymin <- max(round(min(merged.games.plays.tracking2018play392$x, na.rm = TRUE) - 10, -1), 0)
ymax <- min(round(max(merged.games.plays.tracking2018play392$x, na.rm = TRUE) + 10, -1), 120)
df.hash <- expand.grid(x = c(0, 23.36667, 29.96667, xmax), y = (10:110))
df.hash <- df.hash %>% filter(!(floor(y %% 5) == 0))
df.hash <- df.hash %>% filter(y < ymax, y > ymin)

animate.play2 <- ggplot() +
  scale_size_manual(values = c(6, 4, 6), guide = "none") + 
  scale_shape_manual(values = c(21, 16, 21), guide = "none") +
  scale_fill_manual(values = c("#e31837", "#654321", "#002244"), guide = "none") + 
  scale_colour_manual(values = c("black", "#654321", "#c60c30"), guide = "none") + 
  annotate("text", x = df.hash$x[df.hash$x < 55/2], 
           y = df.hash$y[df.hash$x < 55/2], label = "_", hjust = 0, vjust = -0.2) + 
  annotate("text", x = df.hash$x[df.hash$x > 55/2], 
           y = df.hash$y[df.hash$x > 55/2], label = "_", hjust = 1, vjust = -0.2) + 
  annotate("segment", x = xmin, 
           y = seq(max(10, ymin), min(ymax, 110), by = 5), 
           xend =  xmax, 
           yend = seq(max(10, ymin), min(ymax, 110), by = 5)) + 
  annotate("text", x = rep(hash.left, 11), y = seq(10, 110, by = 10), 
                    label = c("G   ", seq(10, 50, by = 10), rev(seq(10, 40, by = 10)), "   G"), 
                    angle = 270, size = 4) + 
  annotate("text", x = rep((xmax - hash.left), 11), y = seq(10, 110, by = 10), 
           label = c("   G", seq(10, 50, by = 10), rev(seq(10, 40, by = 10)), "G   "), 
           angle = 90, size = 4) + 
  annotate("segment", x = c(xmin, xmin, xmax, xmax), 
           y = c(ymin, ymax, ymax, ymin), 
           xend = c(xmin, xmax, xmax, xmin), 
           yend = c(ymax, ymax, ymin, ymin), colour = "black") + 
  geom_point(data = merged.games.plays.tracking2018play392, aes(x = (xmax-y), y = x, shape = team,
                                 fill = team, group = nflId, size = team, colour = team), alpha = 0.7) + 
  geom_text(data = merged.games.plays.tracking2018play392, aes(x = (xmax-y), y = x, label = jerseyNumber), colour = "white", 
            vjust = 0.36, size = 3.5) + 
  ylim(ymin, ymax) + 
  coord_fixed()

animate.play2
```

Choose specific play to visualize (touchback).

```{r}
merged.games.plays.tracking2019play3603 <- filter(merged.games.plays.tracking2019, playId == 3603
 & gameId == 2019090808)
```

Visualize the play.


```{r}
## General field boundaries
xmin <- 0
xmax <- 160/3
hash.right <- 38.35
hash.left <- 12
hash.width <- 3.3


## Specific boundaries for a given play
ymin <- max(round(min(merged.games.plays.tracking2019play3603$x, na.rm = TRUE) - 10, -1), 0)
ymax <- min(round(max(merged.games.plays.tracking2019play3603$x, na.rm = TRUE) + 10, -1), 120)
df.hash <- expand.grid(x = c(0, 23.36667, 29.96667, xmax), y = (10:110))
df.hash <- df.hash %>% filter(!(floor(y %% 5) == 0))
df.hash <- df.hash %>% filter(y < ymax, y > ymin)

animate.play3 <- ggplot() +
  scale_size_manual(values = c(6, 4, 6), guide = "none") + 
  scale_shape_manual(values = c(21, 16, 21), guide = "none") +
  scale_fill_manual(values = c("#e31837", "#654321", "#002244"), guide = "none") + 
  scale_colour_manual(values = c("black", "#654321", "#c60c30"), guide = "none") + 
  annotate("text", x = df.hash$x[df.hash$x < 55/2], 
           y = df.hash$y[df.hash$x < 55/2], label = "_", hjust = 0, vjust = -0.2) + 
  annotate("text", x = df.hash$x[df.hash$x > 55/2], 
           y = df.hash$y[df.hash$x > 55/2], label = "_", hjust = 1, vjust = -0.2) + 
  annotate("segment", x = xmin, 
           y = seq(max(10, ymin), min(ymax, 110), by = 5), 
           xend =  xmax, 
           yend = seq(max(10, ymin), min(ymax, 110), by = 5)) + 
  annotate("text", x = rep(hash.left, 11), y = seq(10, 110, by = 10), 
                    label = c("G   ", seq(10, 50, by = 10), rev(seq(10, 40, by = 10)), "   G"), 
                    angle = 270, size = 4) + 
  annotate("text", x = rep((xmax - hash.left), 11), y = seq(10, 110, by = 10), 
           label = c("   G", seq(10, 50, by = 10), rev(seq(10, 40, by = 10)), "G   "), 
           angle = 90, size = 4) + 
  annotate("segment", x = c(xmin, xmin, xmax, xmax), 
           y = c(ymin, ymax, ymax, ymin), 
           xend = c(xmin, xmax, xmax, xmin), 
           yend = c(ymax, ymax, ymin, ymin), colour = "black") + 
  geom_point(data = merged.games.plays.tracking2019play3603, aes(x = (xmax-y), y = x, shape = team,
                                 fill = team, group = nflId, size = team, colour = team), alpha = 0.7) + 
  geom_text(data = merged.games.plays.tracking2019play3603, aes(x = (xmax-y), y = x, label = jerseyNumber), colour = "white", 
            vjust = 0.36, size = 3.5) + 
  ylim(ymin, ymax) + 
  coord_fixed()

animate.play3
```


Choose specific play to visualize (kickoff returned for a touchdown).

```{r}
merged.games.plays.tracking.2020play1853 <- filter(merged.games.plays.tracking.all, playId == 1853
 & gameId == 2020111600)
```

Visualize the play.


```{r}
## General field boundaries
xmin <- 0
xmax <- 160/3
hash.right <- 38.35
hash.left <- 12
hash.width <- 3.3


## Specific boundaries for a given play
ymin <- max(round(min(merged.games.plays.tracking.2020play1853$x, na.rm = TRUE) - 10, -1), 0)
ymax <- min(round(max(merged.games.plays.tracking.2020play1853$x, na.rm = TRUE) + 10, -1), 120)
df.hash <- expand.grid(x = c(0, 23.36667, 29.96667, xmax), y = (10:110))
df.hash <- df.hash %>% filter(!(floor(y %% 5) == 0))
df.hash <- df.hash %>% filter(y < ymax, y > ymin)

animate.play4 <- ggplot() +
  scale_size_manual(values = c(6, 4, 6), guide = "none") + 
  scale_shape_manual(values = c(21, 16, 21), guide = "none") +
  scale_fill_manual(values = c("#e31837", "#654321", "#002244"), guide = "none") + 
  scale_colour_manual(values = c("black", "#654321", "#c60c30"), guide = "none") + 
  annotate("text", x = df.hash$x[df.hash$x < 55/2], 
           y = df.hash$y[df.hash$x < 55/2], label = "_", hjust = 0, vjust = -0.2) + 
  annotate("text", x = df.hash$x[df.hash$x > 55/2], 
           y = df.hash$y[df.hash$x > 55/2], label = "_", hjust = 1, vjust = -0.2) + 
  annotate("segment", x = xmin, 
           y = seq(max(10, ymin), min(ymax, 110), by = 5), 
           xend =  xmax, 
           yend = seq(max(10, ymin), min(ymax, 110), by = 5)) + 
  annotate("text", x = rep(hash.left, 11), y = seq(10, 110, by = 10), 
                    label = c("G   ", seq(10, 50, by = 10), rev(seq(10, 40, by = 10)), "   G"), 
                    angle = 270, size = 4) + 
  annotate("text", x = rep((xmax - hash.left), 11), y = seq(10, 110, by = 10), 
           label = c("   G", seq(10, 50, by = 10), rev(seq(10, 40, by = 10)), "G   "), 
           angle = 90, size = 4) + 
  annotate("segment", x = c(xmin, xmin, xmax, xmax), 
           y = c(ymin, ymax, ymax, ymin), 
           xend = c(xmin, xmax, xmax, xmin), 
           yend = c(ymax, ymax, ymin, ymin), colour = "black") + 
  geom_point(data = merged.games.plays.tracking.2020play1853, aes(x = (xmax-y), y = x, shape = team,
                                 fill = team, group = nflId, size = team, colour = team), alpha = 0.7) + 
  geom_text(data = merged.games.plays.tracking.2020play1853, aes(x = (xmax-y), y = x, label = jerseyNumber), colour = "white", 
            vjust = 0.36, size = 3.5) + 
  ylim(ymin, ymax) + 
  coord_fixed()

animate.play4
```